---
title: "Quantitative_Assignment_2"
author: "Lasya Madineni"
output:
  pdf_document:
    latex_engine: xelatex
---



# Linear Programming Problem Formulation

## Decision Variables:

- Let:
  - \( L_1 \): Number of units of the Large size product from Plant 1
  - \( M_1 \): Number of units of the Medium size product from Plant 1
  - \( S_1 \): Number of units of the Small size product from Plant 1
  - \( L_2 \): Number of units of the Large size product from Plant 2
  - \( M_2 \): Number of units of the Medium size product from Plant 2
  - \( S_2 \): Number of units of the Small size product from Plant 2
  - \( L_3 \): Number of units of the Large size product from Plant 3
  - \( M_3 \): Number of units of the Medium size product from Plant 3
  - \( S_3 \): Number of units of the Small size product from Plant 3

## Objective Function:

Maximize the total profit:
\[ Z_{\text{max}} = 420(L_1 + L_2 + L_3) + 360(M_1 + M_2 + M_3) + 300(S_1 + S_2 + S_3) \]

## Constraints:

- Production Constraints:
  - \( L_1 + M_1 + S_1 \leq 750 \)
  - \( L_2 + M_2 + S_2 \leq 900 \)
  - \( L_3 + M_3 + S_3 \leq 450 \)
  - \( 20L_1 + 15M_1 + 12S_1 \leq 13,000 \)
  - \( 20L_2 + 15M_2 + 12S_2 \leq 12,000 \)
  - \( 20L_3 + 15M_3 + 12S_3 \leq 5,000 \)
  - \( L_1 + L_2 + L_3 \leq 900 \)
  - \( M_1 + M_2 + M_3 \leq 1,200 \)
  - \( S_1 + S_2 + S_3 \leq 750 \)

- Non-negativity Constraints:
  - \( L_1, L_2, L_3, M_1, M_2, M_3, S_1, S_2, S_3 \geq 0 \)

This formulation aims to optimize the production process across the three plants while adhering to the specified constraints.

\[
\begin{aligned}
& L1 + M1 + S1 + 0L2 + 0M2 + 0S2 + 0L3 + 0M3 + 0S3 \leq 750 \\
& 0L1 + 0M1 + 0S1 + L2 + M2 + S2 + 0L3 + 0M3 + 0S3 \leq 900 \\
& 0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + L3 + M3 + S3 \leq 450 \\
& 20L1 + 15M1 + 12S1 + 0L2 + 0M2 + 0S2 + 0L3 + 0M3 + 0S3 \leq 13000 \\
& 0L1 + 0M1 + 0S1 + 20L2 + 15M2 + 12S2 + 0L3 + 0M3 + 0S3 \leq 12000 \\
& 0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + 20L3 + 15M3 + 12S3 \leq 5000 \\
& L1 + 0M1 + 0S1 + L2 + 0M2 + 0S2 + L3 + 0M3 + 0S3 \leq 900 \\
& 0L1 + M1 + 0S1 + 0L2 + M2 + 0S2 + 0L3 + M3 + 0S3 \leq 1200 \\
& 0L1 + 0M1 + S1 + 0L2 + 0M2 + S2 + 0L3 + 0M3 + S3 \leq 750 \\
\end{aligned}
\]


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# If not already installed, install and load the lpSolve library

```{r lpsolve}

if (!require(lpSolve)) {
  install.packages("lpSolve")
  library(lpSolve)
}
```

# Coefficients of the objective function for all three plants

```{r objective}
obj_coef <- rep(c(420, 360, 300), 3)

```

# Limitations on storage space and production capacity

```{r storage_production}

const_coef <- rbind(
# Production capacity
c(1, 1, 1, 0, 0, 0, 0, 0, 0),
c(0, 0, 0, 1, 1, 1, 0, 0, 0),
c(0, 0, 0, 0, 0, 0, 1, 1, 1),
# Storage space
c(20, 15, 12, 0, 0, 0, 0, 0, 0),
c(0, 0, 0, 20, 15, 12, 0, 0, 0),
c(0, 0, 0, 0, 0, 0, 20, 15, 12)
)
```


# Values on the right-hand side for storage space and production capacity restrictions
```{r constant}
rhs_const <- c(750, 900, 450, 13000, 12000, 5000)
```


# Incorporate sales forecasts as additional constraints
```{r sales_forecasrs_constraints}
sales_forecast_constraints <- matrix(0, 3, 9)
sales_forecast_constraints[1, c(1,4,7)] <- 1
sales_forecast_constraints[2, c(2,5,8)] <- 1
sales_forecast_constraints[3, c(3,6,9)] <- 1
```

# Consolidate the constraints
```{r combine_const}
const_coef <- rbind(const_coef, sales_forecast_constraints)
rhs_const <- c(rhs_const, c(900, 1200, 750))
```

# Directions of constraints (all are '<=')
```{r const_dir}
const_dir <- rep("<=", nrow(const_coef))
```

# Resolve the LP model
```{r lp_sol}
lp_solution <- lp(direction = "max", objective.in = obj_coef, const.mat = const_coef, const.dir = const_dir, const.rhs = rhs_const, all.int = TRUE)
```

# Print
```{r print}
print(lp_solution)

```


# Retrieve solution values

```{r prod_plan}
production_plan <- lp_solution$solution


# Format and display the production plan for each plant

for (plant in 1:3) {
start_idx <- (plant - 1) * 3 + 1
end_idx <- start_idx + 2
plant_production <- production_plan[start_idx:end_idx]
cat(sprintf("Plant %d:\n", plant))
cat(sprintf(" Large: %d units\n", plant_production[1]))
cat(sprintf(" Medium: %d units\n", plant_production[2]))
cat(sprintf(" Small: %d units\n\n", plant_production[3]))
}
```

